cvx-flats

### 2020-04-09 neurips-20/cvx-flats.tex
First pass on proof of main theorem.  Still a bit hazy.

\begin{definition}[Consistent]
	Suppose $f_m : \X \to \R$ is the hypothesis function learned by minimizing empirical training loss over $m$ labeled examples.
	For a distribution $D$ over input and label spaces $\X \times\Y$, we say $\{f_m\}$ is \emph{consistent} with respect to a loss $L: \R \times \Y \to \reals_+$ if
	\begin{align*}
	\E_D L(f_m(X), Y) &\to \inf_f \E_D L(f(X), Y)~.~
	\end{align*}
	Moreover, we say a loss and link $(L,\psi)$ are consistent with respect to a discrete loss $\ell$ if 
	\begin{align*}
		\E_D L(f_m(X), Y) \to \inf_f \E_D L(f(X), Y) &\implies \E_D \ell(\psi \circ f_m(X), Y) \to \inf_f \E_D \ell(\psi \circ f(X), Y)~.~
	\end{align*}
\end{definition}

@@

Similarly, we can consider the notion of \emph{identifiability} for vector-valued properties.

\begin{definition}[$d$-identifiable, Identification Complexity]
	A vector-valued property $\Gamma: \simplex \to \reals^d$ is $d$-identifiable if it holds that $p \in \Gamma_u \iff V(u, p) = \vec 0 \in \reals^d$.
	Like elicitation complexity, we define the \emph{identification complexity} of a property $\Gamma$ as the minimum dimension $d$ such that there is a link $\psi$ such that $\Gamma = \psi \circ \hat \Gamma$ and $\hat \Gamma$ is $d$-identifiable.
\end{definition}

@@

\section{Implications of property elicitation for consistency}
\begin{proposition}\label{prop:consistent-implies-indir-elic}
	If a surrogate and link pair $(L, \psi)$ is consistent with respect to $\ell$, then $L$ indirectly elicits $\gamma := \prop{\ell}$.
\end{proposition}
\begin{proof}
Since consistency must hold for all $D$ over $\X \times \Y$, let us focus on distributions $D$ such that the probability of any $x \in \X$ is a point mass, allowing us to focus on conditional distributions $p \in \simplex$ such that $p = Pr[Y = y | X = x]$, allowing us to be agnostic to $\X$.
Now, instead of focusing on sequences of optimal hypotheses, we can focus on optimal reports when given $x \in \X$ if the hypothesis class $\F$ is sufficiently rich (i.e. the optimal hypothesis is in the hypothesis class for all $x \in \X$.)
 \jessie{FINISH}
\end{proof}

@@

\section{Characteristics of convex surrogates}\label{sec:char-convex}
When one restricts their focus to \emph{convex} surrogates, we know that all minima are global, and we then have $\nabla L(u, y) = \vec 0$ if and only if $u$ minimizes $L$ in its first component.

\begin{theorem}
	Suppose the property $\Gamma$ is elicited by the convex loss $L : \reals^d \times \Y \to \reals_+$ such that the infimum $L(\cdot)_y$ is attained for all $y \in \Y$.
	Then for all $p \in \simplex$, we have $u \in \Gamma(p) \iff \vec 0 \in \nabla_{u'} \inprod{p}{L(u')}$.
\end{theorem}
\begin{proof}[Proof sketch]
	Since $L$ is convex, we know that all minimizers are global. 
	This implies that $\vec 0 \in \nabla_{u} L(u; p) \iff u \in \arginf_{u'} L(u';p)$, and the latter is true if and only if $u \in \Gamma(p)$ by definition of elicitation.
\end{proof}

In a large sense, this lets us conclude that $d$-convex elicitable properties are also $d$-identifiable, taking the identification function $V(\cdot;p)$ to be $\nabla_u L(\cdot; p)$ for all $p \in \simplex$.
\begin{corollary}
	Suppose the property $\Gamma$ is elicited by the convex loss $L : \reals^d \times \Y \to \reals_+$ such that the infimum $L(\cdot, y)$ is attained for all $y \in \Y$.
	Then $\Gamma$ is $d$-identifiable.
\end{corollary}

@@

\begin{theorem}
	Suppose we are given a discrete loss $\ell:\R \times \Y \to \reals_+$ eliciting $\gamma:= \prop{\ell}$.
	Fix $p \in \simplex$, and let $r$ be the report such that $p \in \gamma_r$.  
	If there is not $(n - d - 1)$-dimensional flat $F$ such that $F \subseteq \gamma_r$ and $p \in F$, then there is no surrogate loss $L : \reals^d \times \Y \to \reals_+$ and link $\psi:\reals^d \to \R$ such that $(L, \psi)$ is calibrated with respect to $\ell$.
	Moreover, this must hold for all $p \in \simplex$ and $r \in \gamma(p)$.
\end{theorem}
\begin{proof}[Proof intuition]
	\jessie{TODO}
\end{proof}

@@

>>>>

\begin{definition}[Consistent]
	Suppose $f_m : \X \to \R$ is the hypothesis function learned by minimizing empirical training loss over $m$ labeled examples.
	For all distributions $D$ over input and label spaces $\X \times\Y$, we say a loss and link $(L,\psi)$ are consistent with respect to an original loss $\ell$ if 
	\begin{align*}
		\E_D L(f_m(X), Y) \to \inf_f \E_D L(f(X), Y) &\implies \E_D \ell(\psi \circ f_m(X), Y) \to \inf_f \E_D \ell(\psi \circ f(X), Y)~.~
	\end{align*}
\end{definition}

@@

\iffalse
Similarly, we can consider the notion of \emph{identifiability} for vector-valued properties.

\begin{definition}[$d$-identifiable, Identification Complexity]
	A vector-valued property $\Gamma: \simplex \to \reals^d$ is $d$-identifiable if it holds that $p \in \Gamma_u \iff V(u, p) = \vec 0 \in \reals^d$.
	Like elicitation complexity, we define the \emph{identification complexity} of a property $\Gamma$ as the minimum dimension $d$ such that there is a link $\psi$ such that $\Gamma = \psi \circ \hat \Gamma$ and $\hat \Gamma$ is $d$-identifiable.
\end{definition}
\fi

@@

\section{Characteristics of Consistent Convex Surrogates}\label{sec:char-convex}
\jessie{Intro text}
\begin{proposition}\label{prop:consistent-implies-indir-elic}
	If a surrogate and link pair $(L, \psi)$ is consistent with respect to $\ell$, then $L$ indirectly elicits $\gamma := \prop{\ell}$.
\end{proposition}
\begin{proof}
Since consistency must hold for all $D$ over $\X \times \Y$, let us focus on distributions $D$ such that the probability of any $x \in \X$ is a point mass, allowing us to focus on conditional distributions $p \in \simplex$ such that $p = Pr[Y = y | X = x]$, allowing us to be agnostic to $\X$.
Now, instead of focusing on sequences of optimal hypotheses, we can focus on optimal reports when given $x \in \X$ if the hypothesis class $\F$ is sufficiently rich (i.e. the optimal hypothesis is in the hypothesis class for all $x \in \X$.)
 \jessie{FINISH}
\end{proof}

@@

When one restricts their focus to \emph{convex} surrogates, we know that all minima are global, and we then have $\vec 0  \in \partial L(u, y)$ if and only if $u$ minimizes $L$ in its first component.

%\begin{theorem}
%	Suppose the property $\Gamma$ is elicited by the differentiable, convex loss $L : \reals^d \times \Y \to \reals_+$ such that the infimum $L(\cdot, y)$ is attained for all $y \in \Y$.
%	Then for all $p \in \simplex$, we have $u \in \Gamma(p) \iff \vec 0 \in \nabla_{u'} \inprod{p}{L(u')}$.
%\end{theorem}
%\begin{proof}[Proof sketch]
%	Since $L$ is convex, we know that all minimizers are global. 
%	This implies that $\vec 0 \in \nabla_{u} L(u; p) \iff u \in \arginf_{u'} L(u';p)$, and the latter is true if and only if $u \in \Gamma(p)$ by definition of elicitation.
%\end{proof}
%
%In a large sense, this lets us conclude that $d$-convex elicitable properties are also $d$-identifiable, taking the identification function $V(\cdot;p)$ to be $\nabla_u L(\cdot; p)$ for all $p \in \simplex$.
%\begin{corollary}
%	Suppose the property $\Gamma$ is elicited by the convex loss $L : \reals^d \times \Y \to \reals_+$ such that the infimum $L(\cdot, y)$ is attained for all $y \in \Y$.
%	Then $\Gamma$ is $d$-identifiable.
%\end{corollary}

@@

\begin{theorem}\label{thm:cvx-flats}
	Suppose we are given an elicitable property $\gamma$.
	Fix $p \in \simplex$, and let $r$ be a report such that $p \in \gamma_r$.  
	If there is not $(n - d - 1)$-dimensional flat $F$ such that $F \subseteq \gamma_r$ and $p \in F$, then there is no convex surrogate loss $L : \reals^d \times \Y \to \reals_+$ that indirectly elicits $\gamma$.
	Moreover, this must hold for all $p \in \simplex$ and $r \in \gamma(p)$.
\end{theorem}
\begin{proof}[Proof intuition]
	We prove the contrapositive; if there is a surrogate loss $L$ indirectly eliciting $\gamma$, then for all $p \in \simplex$ and $r \in \R$ such that $r \in \gamma(p)$, there is an $(n-d-1)$-dimensional flat $F$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$.

	Suppose we have $L:\reals^d \times \Y \to \reals$ eliciting $\Gamma$ and link $\psi : \reals^d \to \R$ such that for all $u \in \reals^d$, we have $\Gamma_u \subseteq \gamma_{\psi(u)}$.
	Fix any $p \in \simplex$, and take any $r \in \gamma(p)$.
	We argue that there must be some $u \in \reals^d \in \Gamma(p)$ so that $r = \psi(u)$, since $\Gamma(p) = \emptyset$ otherwise.
	We know that $r \in \gamma(p) \iff u \in \Gamma(p) \iff \vec 0 \in \partial L(u; p) \iff \vec 0 \in \oplus_y p_y \partial L(u,y)$, where $\oplus$ denotes the nonnegative weighted Minkowski sum by convex analysis results \jessie{And mild assumptions on the loss?}.

	The weighted Minkowski sum is in $d$-dimensional space as $L$ is defined on $\reals^d$.
	We claim there exists a matrix $W \in \reals^{d+1 \times n}$ defining the flat $F = \{q \in \reals^n : Wq = \vec 0 \}$; $F$ is then $(n-(d+1))$-dimensional by the rank-nullity theorem, so we just need to show such a $W$ also keeps $F \cap \simplex \subseteq \gamma_r$.
	As $\Gamma_u \subseteq \gamma_r$, it is sufficient to show $F \cap \simplex \subset \Gamma_u$.

	Construct $W$ by taking each column to describe $\partial L(u, y)$ for each $y \in \Y$.
	Since $\partial L(u, y)$ is $d$-dimensional, the affine subspace containing each subgradient can be described with $d+1$ constraints. \jessie{??}
	We can see that $p \in F$ since $\vec 0 \in \oplus_y p_y \partial L(u, y) \implies Wp = \vec 0$ by construction of $W$.
	Moreover, if there was a $\hat q \in \simplex \setminus \gamma_r$ such that $W \hat q = \vec 0$, then we would contradict the indirect elicitation of $\gamma$ since we would not have $\Gamma_{u} \subseteq \gamma_{\psi(u)}$ since we would have $\hat q \in \Gamma_u$ but $\hat q \not \in \gamma_{\psi(u)}$.
	Therefore,  we have an $(n - d - 1)$-dimensional flat $F$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$ for all $p \in \simplex$ and $r \in \gamma(p)$.
\end{proof}

@@

\section{Continuous Properties}
Theorem~\ref{thm:} allows us to address one major open question from~\cite{frongillo2015elicitation}: \emph{what are lower bounds on convex elicitation complexity?}
